---
title: "BCI - Eye State"
output: html_document
date: "2024-05-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## EYE State 
It is quite important in the applications of medicine; for example, the disabled people are able to interact better by utilizing this way, and this is achieved by founding a method that acts like a bridge between computer and human brain (Ahmadi et al). 


```{r cars}
# Load Libraries
library(rio)
library(rstatix)
library(tidyverse)
library(dplyr)
library(FactoMineR)
library(factoextra)
library(corrplot)

# import data
eeg <- import("EEG/EEG Eye State.arff")
```

# Coding of EEG cariables

- AF3: Frontal lobe, left hemisphere
- F7: Frontal lobe, left hemisphere, near the temple
- F3: Frontal lobe, left hemisphere
- FC5: Frontal lobe, left hemisphere, near the center
- T7: Temporal lobe, left hemisphere
- P7: Parietal lobe, left hemisphere
- O1: Occipital lobe, left hemisphere
- O2: Occipital lobe, right hemisphere
- P8: Parietal lobe, right hemisphere
- T8: Temporal lobe, right hemisphere
- FC6: Frontal lobe, right hemisphere, near the center
- F4: Frontal lobe, right hemisphere
- F8: Frontal lobe, right hemisphere, near the temple
- AF4: Frontal lobe, right hemisphere

```{r}
## DATA ANALYSIS
# Eye Closed
eye_closed <- eeg[eeg$eyeDetection == 1, ]
# Eye Open
eye_open <- eeg[eeg$eyeDetection == 0, ]

# dropping the "eye detection" column
eye_closed <- eye_closed[, -15]
eye_open <- eye_open[, -15]

## get STATS
eeg_closed_stats <- get_summary_stats(eye_closed)
eye_open_stats <- get_summary_stats(eye_open)

# store in a data frame
df_eeg_closed_stats <- as.data.frame(eeg_closed_stats)
df_eye_open_stats <- as.data.frame(eye_open_stats)
```

```{r pressure}
# Overall dataset
# dropping the "eye detection" column
eeg_variables <- eeg[, -15]

# correlation matrix
corr_eeg <- cor(eeg_variables)
corr_eeg <- round(corr_eeg, 2)

corrplot(corr_eeg)

```

```{r pressure}
## PCA plotting
# - Dimensionality reduction
# - Data visualization
# - Feature engineering

eeg_prcomp <- prcomp(eeg_variables, scale=TRUE)
df_eeg_prcomp <- as.data.frame(eeg_prcomp$x)
df_eeg_prcomp$label <- eeg$eyeDetection

plot(eeg_prcomp_x$PC1,
     eeg_prcomp_x$AF3,
     col = eeg_prcomp_x$label,
     pch = 19,
     main="Figure 1: 2D Scatterplot",
     log = "y",
     xlab="PC1",
     ylab="AF3")
legend("bottomright",
       legend = levels(eeg_prcomp_x$label),
       col = seq_along(levels(eeg_prcomp_x$label)),
       pch = 19)

```


```{r}
# scree plot of variance
fviz_eig (eeg_prcomp, addlabels = TRUE)
# # Biplot
fviz_pca_biplot(eeg_prcomp)
# eeg_prcomp_x$label <- eeg$eyeDetection
# #Biplot with colored groups
fviz_pca_biplot(eeg_prcomp, label = "var", habillage = df_eeg_prcomp$label)
# install.packages("caret")
# View(eeg)
```
# Machine Learning plots
